String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-06 2023-1-19/sections/implementation.tex
experiments

We implement the algorithms in Section~\ref{sec:algorithm} on top of OSTRICH \cite{ostrich,atva2020}. Then we do experiments on two benchmark suites, to evaluate the efficacy of our approach.  In the sequel, we first describe the two benchmark suites, then present the settings and results of the experiments.

@@

Our experiments utilize two benchmark suites, namely, AutomatArk and RegCoLe. In total, there are 49,379 instances in the two benchmark suites.

@@

\paragraph*{RegCoLe benchmark suite.} There are 40,628 RECL constraints in the RegCoLe suite.  These constraints are generated by extracting regular expressions with counting operators from \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form 
$$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$$   
where $\Sigma^*(<+ >+'+''+\&)\Sigma^*$ is a regular expression that specifies at least one occurrence of special characters $<$, $>$, $'$, $''$, or $\&$, and $\overline{e'}$ denotes the complement of a regular expression $e'$. We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

@@

\subsection{Experiments}

We have evaluated OstrichCEA compared to five other prominent string solvers currently available. We evaluate the solvers by directly comparing the number of cases correctly solved, the average time taken with and without timeouts, and the total count of soundness errors and program crashes. One of these solvers is CVC5\cite{cvc5}, a general SMT solver that uses algebraic reasoning to handle strings and regular expressions and is the winner of SMT-COMP 2022\cite{smt-comp}. Another solver, Z3str3\cite{z3str3}, is the most recent addition we can get to the Z3-str family and utilizes a word equation reduction approach to reason about regular expressions. Z3str3RE\cite{z3str3re} is a variant of Z3str3 that incorporates length-aware algorithms and heuristics. Z3seq\cite{z3seq} is a sequence solver which uses a novel derivative theory for solving extended regular expressions. Z3-Trau\cite{z3trau} is the Z3 version of trau\cite{trau} that employs a flat automata-based approach, incorporating both under- and over-approximations. Ostrich\cite{ostrich} is the tool we extend which uses automaton to model the semantics of string functions and regular memberships. We used the 1.0.5 binary version of CVC5, commit 59e9c87 of Z3str3, last version of Z3str3RE, 4.8.9 binary version of Z3Seq, commit 1628747 of Z3-Trau and commit 8297d8d of Ostrich. Z3-Trau does not support \verb|re.diff|. All other solvers support all syntax sugars listed in SMT-LIB standard\cite{smt_lib}. We omitted Z3str4\cite{z3str4} because the provided reproduction package link is wrong. All experiments are conducted on CentOS Stream release 8 with 12 Intel(R) Xeon(R) Platinum 8269CY CPU T 3.10GHz processors and 190 GB memory. We used Zaligvinder\cite{zaligvinder_2021} framework and set the timeout to 60 seconds.

\subsection{Overall Evaluation}
In Fig.\ref{fig:cactus_all}, the cactus plot illustrates the cumulative time each solver takes for all cases in ascending order of runtime. Solvers located towards the right and lower portion of the plot indicate better performance. \newline
Table \ref{tab:results_all} summarizes the results that demonstrate OstrichCEA's superior performance, solving the most significant number of instances and outperforming most competing solvers. Including timeouts, OstrichCEA is \textbf{1.52}\mult{} faster than Z3Seq, \textbf{2.23}\mult{} faster than Ostrich, \textbf{2.26}\mult{} faster than Z3-Trau, \textbf{3.08}\mult{} faster than Z3str3, \textbf{3.11}\mult{} faster than CVC5 and close to Z3str3RE with \%2 speed loss . Note that CVC5\cite{cvc5} yielded 5370 timeouts(11\% of all instances), and Z3str3\cite{z3str3} yielded 6139 timeouts(12\% of all instances), which is much more than other solvers. Both CVC5 and Z3str3 are DPLL(T)-based solvers. It seems that almost 10\% of the benchmarks we used seem unsuitable for them. Z3-trau\cite{z3trau} yielded 21,152 unknowns because it does not support \verb|re.diff|. Z3-trau\cite{z3trau} yielded 6673 crashes (13\% of the instances) and 1233 soundness errors (2\% of the instances), which is a large portion. Z3-based solvers Z3str3 yielded 38 soundness errors, Z3seq\cite{z3seq} yielded 51 soundness errors and Z3str3RE\cite{z3str3re} yielded 39 soundness error. Most of them are due to the mistake formalization of the backslash character. OstrichCEA resulted 5 unknowns because it reached the maximum threshold of limited memory 2GB. More details of the results of each benchmark are shown in \ref{appendix:experiential_results}.
\begin{figure}[h]
  \centering
  \import{figures}{cactus_plot_all.tex}
  \caption{Cactus plot summarizing performance on all benchmarks.}
  \label{fig:cactus_all}
\end{figure}
\begin{table}
  \import{tables}{table_all.tex}
  \caption{Total results of string solvers on all benchmarks. OstrichCEA solved the most benchmarks in the second shortest time.}
  \label{tab:results_all}
\end{table}

\subsection{Analysis of Individual Heuristics}
In order to demonstrate the efficacy of the individual heuristics outlined in Section \ref{sec:algorithm} and incorporated into OstrichCEA, we assessed various tool configurations in which one or more heuristics were disabled. Figure \ref{fig:cactus_heuristics} and Table \ref{tab:results_heuristics} display the outcomes. The "OstrichCEA" plot line represents the tool's performance when all heuristics are enabled, while the "All heuristics off" line represents performance when all heuristics are disabled. The remaining plot lines exhibit performance with only the named heuristic disabled while all others are enabled. From the plots and table, it is evident that OstrichCEA functions most effectively when all heuristics are enabled. In average time with timeout times, OstrichCEA performs \textbf{1.78}$\times$ faster than employing none of our heuristics, \textbf{1.37}$\times$ faster than turning off simplification, \textbf{1.15}$\times$ faster than truing off under-approximation.
\begin{figure}
  \subimport{figures/}{cactus_plot_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics using a cactus plot.}
  \label{fig:cactus_heuristics}
\end{figure}
\begin{table}
  \subimport{tables}{table_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics.}
  \label{tab:results_heuristics}
\end{table}

%\end{document}

>>>>

We implement the algorithms in Section~\ref{sec:algorithm} on top of OSTRICH~\cite{ostrich,atva2020}, resulting into a version of OSTRICH called OSTRICH$^{\rm RECL}$.
Then we do experiments on two benchmark suites, to evaluate the performance of OSTRICH$^{\rm RECL}$.  In the sequel, we first describe the two benchmark suites, then present the settings and results of the experiments.

@@

Our experiments utilize two benchmark suites, namely, AutomatArk and RegCoL. In total, there are 49,379 instances in the two benchmark suites. All the benchmark instances are in SMTLIB2 format.

@@

\paragraph*{RegCoL benchmark suite.} There are 40,628 RECL constraints in the RegCoL suite.  These constraints are generated by extracting regular expressions with counting operators from \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form 
$$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$$   
where $\Sigma^*(<+ >+'+''+\&)\Sigma^*$ is a regular expression that specifies at least one occurrence of special characters $<$, $>$, $'$, $''$, or $\&$, and $\overline{e'}$ denotes the complement of a regular expression $e'$. We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

@@

\subsection{Experiment results}

We evaluated the performance of OSTRICH$^{\rm RECL}$ against the main-stream string constraint solvers, including CVC5 (version 1.0.5) \cite{cvc5}, Z3seq (version 4.8.9) \cite{z3seq}, Z3-Trau (github commit 1628747) \cite{z3trau}, Z3str3 (github commit 59e9c87) \cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH (github commit 8297d8d) \cite{atva2020}. All experiments are conducted on CentOS Stream release 8 with 12 Intel(R) Xeon(R) Platinum 8269CY 3.10GHz CPU cores and 190 GB memory. We use the Zaligvinder framework \cite{zaligvinder_2021} to execute the experiments, with the timeout limit set as 60 seconds per instance.

The experiments are designed to answer the following two research questions. 
\begin{description}
\item[Q1] Is OSTRICH$^{\rm RECL}$ more efficient to solve RECL constraints than the state-of-the-art string constraint solvers ?

\item[Q2] Do the size-reduction and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
\end{description}

Q1 is answered by running OSTRICH$^{\rm RECL}$ and the aforementioned solvers on AutomatArk and RegCoL. The experimental results can be found in Table~\ref{tab:results_all}. From Table~\ref{tab:results_all}, we can see that OSTRICH$^{\rm RECL}$ solves 49,128 instances, which is 5,127/2,315/25,369/12,808/1,721/1,756 more than the number of instances solved by CVC5/Z3seq/Z3-Trau/Z3str3/Z3str3RE/OSTRICH respectively. 
%    
Moreover, OSTRICH$^{\rm RECL}$ is almost as fast as Z3str3RE, the fastest solver. More specifically, OSTRICH$^{\rm RECL}$ spends 2.51 seconds per instance on average, while Z3str3RE spends 2.46 seconds per instance. Z3-Trau reports ``unknown'' on 21,152 instances, which is mainly attributed to the fact that it does not support \verb|re.diff|, the language difference operator of regular expressions. Furthermore, Z3seq, Z3-Trau, Z3str3, and Z3str3RE have soundness issues, that is, their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.

\zhilin{may add the results for the two benchmark suites separately.}

\begin{table}[ht]
\begin{center}
  \import{tables}{table_all.tex}
\end{center}
  \caption{Experiment results for Q1}
  \label{tab:results_all}
\end{table}

To answer Q2, we compare OSTRICH$^{\rm RECL}$ with its three variants, namely, OSTRICH$^{\rm RECL}_{\rm -SR}$, OSTRICH$^{\rm RECL}_{\rm -UA}$, OSTRICH$^{\rm RECL}_{\rm -SRUA}$, which remove size-reduction, under-approximation, and both from OSTRICH$^{\rm RECL}$ respectively. The experiment results can be found in Table~\ref{tab:results_heuristics}.
From Table~\ref{tab:results_heuristics}, we can see that adding size-reduction techniques help solve \zhilin{xxx} more instances, while adding under-approximation techniques help solve \zhilin{xxx} more instances, while adding both help solve \zhilin{xxx} more instances. Moreover, adding these techniques also decrease the average time spent in solving each instance considerably.

\begin{table}[ht]
\begin{center}
  \subimport{tables}{table_heuristic.tex}
\end{center}
  \caption{Experiment results for Q2}
  \label{tab:results_heuristics}
\end{table}

\zhilin{stopped here}

%%%%%%%%%% original texts by denghang %%%%%%%%%
%%%%%%%%%% original texts by denghang %%%%%%%%%
\hide{
%
We have evaluated OstrichCEA compared to five other prominent string solvers currently available. We evaluate the solvers by directly comparing the number of cases correctly solved, the average time taken with and without timeouts, and the total count of soundness errors and program crashes. One of these solvers is CVC5\cite{cvc5}, a general SMT solver that uses algebraic reasoning to handle strings and regular expressions and is the winner of SMT-COMP 2022\cite{smt-comp}. Another solver, Z3str3\cite{z3str3}, is the most recent addition we can get to the Z3-str family and utilizes a word equation reduction approach to reason about regular expressions. Z3str3RE\cite{z3str3re} is a variant of Z3str3 that incorporates length-aware algorithms and heuristics. Z3seq\cite{z3seq} is a sequence solver which uses a novel derivative theory for solving extended regular expressions. Z3-Trau\cite{z3trau} is the Z3 version of trau\cite{trau} that employs a flat automata-based approach, incorporating both under- and over-approximations. Ostrich\cite{ostrich} is the tool we extend which uses automaton to model the semantics of string functions and regular memberships. We used the 1.0.5 binary version of CVC5, commit 59e9c87 of Z3str3, last version of Z3str3RE, 4.8.9 binary version of Z3Seq, commit 1628747 of Z3-Trau and commit 8297d8d of Ostrich. Z3-Trau does not support \verb|re.diff|. All other solvers support all syntax sugars listed in SMT-LIB standard\cite{smt_lib}. We omitted Z3str4\cite{z3str4} because the provided reproduction package link is wrong. All experiments are conducted on CentOS Stream release 8 with 12 Intel(R) Xeon(R) Platinum 8269CY CPU T 3.10GHz processors and 190 GB memory. We used Zaligvinder\cite{zaligvinder_2021} framework and set the timeout to 60 seconds.

%\subsection{Overall Evaluation}
In Fig.\ref{fig:cactus_all}, the cactus plot illustrates the cumulative time each solver takes for all cases in ascending order of runtime. Solvers located towards the right and lower portion of the plot indicate better performance. \newline
Table \ref{tab:results_all} summarizes the results that demonstrate OstrichCEA's superior performance, solving the most significant number of instances and outperforming most competing solvers. Including timeouts, OstrichCEA is \textbf{1.52}\mult{} faster than Z3Seq, \textbf{2.23}\mult{} faster than Ostrich, \textbf{2.26}\mult{} faster than Z3-Trau, \textbf{3.08}\mult{} faster than Z3str3, \textbf{3.11}\mult{} faster than CVC5 and close to Z3str3RE with \%2 speed loss . Note that CVC5\cite{cvc5} yielded 5370 timeouts(11\% of all instances), and Z3str3\cite{z3str3} yielded 6139 timeouts(12\% of all instances), which is much more than other solvers. Both CVC5 and Z3str3 are DPLL(T)-based solvers. It seems that almost 10\% of the benchmarks we used seem unsuitable for them. Z3-trau\cite{z3trau} yielded 21,152 unknowns because it does not support \verb|re.diff|. Z3-trau\cite{z3trau} yielded 6673 crashes (13\% of the instances) and 1233 soundness errors (2\% of the instances), which is a large portion. Z3-based solvers Z3str3 yielded 38 soundness errors, Z3seq\cite{z3seq} yielded 51 soundness errors and Z3str3RE\cite{z3str3re} yielded 39 soundness error. Most of them are due to the mistake formalization of the backslash character. OstrichCEA resulted 5 unknowns because it reached the maximum threshold of limited memory 2GB. More details of the results of each benchmark are shown in \ref{appendix:experiential_results}.
\begin{figure}[h]
  \centering
  \import{figures}{cactus_plot_all.tex}
  \caption{Cactus plot summarizing performance on all benchmarks.}
  \label{fig:cactus_all}
\end{figure}

%\subsection{Analysis of Individual Heuristics}
In order to demonstrate the efficacy of the individual heuristics outlined in Section \ref{sec:algorithm} and incorporated into OstrichCEA, we assessed various tool configurations in which one or more heuristics were disabled. Figure \ref{fig:cactus_heuristics} and Table \ref{tab:results_heuristics} display the outcomes. The "OstrichCEA" plot line represents the tool's performance when all heuristics are enabled, while the "All heuristics off" line represents performance when all heuristics are disabled. The remaining plot lines exhibit performance with only the named heuristic disabled while all others are enabled. From the plots and table, it is evident that OstrichCEA functions most effectively when all heuristics are enabled. In average time with timeout times, OstrichCEA performs \textbf{1.78}$\times$ faster than employing none of our heuristics, \textbf{1.37}$\times$ faster than turning off simplification, \textbf{1.15}$\times$ faster than truing off under-approximation.
\begin{figure}
  \subimport{figures/}{cactus_plot_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics using a cactus plot.}
  \label{fig:cactus_heuristics}
\end{figure}
\begin{table}
  \subimport{tables}{table_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics.}
  \label{tab:results_heuristics}
\end{table}
}
%%%%%%%%%% original texts by denghang %%%%%%%%%
%%%%%%%%%% original texts by denghang %%%%%%%%%
%\end{document}
